Whose voices are heard?

### 2024-01-08 paper/article.qmd
major editing

---
title: "Whose voices are heard?"
subtitle: "Making sense of framing patterns in protest coverage"
abstract: |
  Existing research indicates that when mainstream news media report about demonstrations, protesters often face delegitimising coverage that focuses on, for example, clashes with the police or the appearance of protesters instead of their message.
  Recent studies have also found, however, that media treatment differs along several features of protest events.
  In this study, I examined the impact of these features through a large-scale automated framing analysis of UK protest events from 1992 to 2017, complemented by data from the Mass Mobilization Project (MMP).
  My findings highlight four main determinants influencing the use of legitimising or delegitimising framing in protest news coverage:
  (1.) Violence: Protests involving violence receive more delegitimising coverage and less legitimising coverage.
  (2.) Protest goals: The objectives of a protest significantly affect the nature of its coverage, although the relationships between frames and goals are complex. Notably, goals play a larger role in legitimising framing.
  (3.) Outlet characteristics: The type and ideology of a news outlet also impact coverage, though to a lesser extent than anticipated.
  (4.) Timing: More recent reports and those published a significant time after the event's onset are more likely to contain legitimising framing.

format:
  aft-pdf:
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
  aft-html: 
    citeproc: false
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
date: today
filters:
  - abstract-section
keywords:
    - "protest paradigm"
    - "media framing"
    - "social movements"
    - "news reporting"
    - "content analysis"
bibliography: references.bib
execute:
  echo: false
  eval: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  fig.path = "plots/"
)
source("../misc/functions.R")
library(tidyverse); theme_set(theme_minimal())
library(kableExtra)
library(texreg)
library(margins)
library(grid)
```

@@

Protest is a fundamental democratic tool for citizens to voice dissent about issues in society.
In modern democracies, however, citizens usually only hear about protest through mass media.
And this hearing consists not only of the audience becoming aware of a protest event, but ideally should also involve them receiving the information to *understand* the underlying issue and the causes for the public dissent.

However, most of the literature concerning protest coverage in news media points to a phenomenon referred to as *protest paradigm* to explain a recurring patterns in news about demonstrations:
the media focus on crime, clashes with police and the spectacle surrounding a protest or the protesters themselves and their looks, demographics or occupation, while neglecting to explain *why* protesters spent time and resources on organising and participating in a street demonstration in the first place.
The *protest paradigm* explains this by presuming a mental catalogue, supposedly common among journalists, which indicates to them "where to look (and where not to look), and informs them about what to discover" [@chanJournalisticParadigmsCivil1984, p. 187].
The *protest paradigm* is thus closely related to the literature on the sociology of newsmaking, which observed that news workers are "routinizing the unexpected" [@tuchmanMakingNewsDoing1973, p. 110] by creating practical and effective procedural rules, routines and mental catalogues of news story themes to cope with the onslaught of new events.

Most of the focus in previous studies has been spent on the dangers of this routinised reporting.
This is due to the normative core of the *protest paradigm*:
by regularly focusing on the action itself, rather than the sometimes more elusive and complicated message [@smithProtestAgendaBuilding2001] reporting shaped by the *protest paradigm* could lead to delegitimisation and marginalisation of protest events, protest movements, or protest as a political resource [@mcleodSocialControlSocial1999].

*Some* evidence for the *protest paradigm* has been found in most studies that analysed news coverage of protests over the years [for a systematic overview see @gruberTrouble2023].
However, more recent work has also asked *how common* delegitimising devices of the *protest paradigm* are, as nobody can deny that the press at least occasionally discusses the socio-political issues protesters try to rise awareness for --- hence *legitimising their actions*.
While it seems fairly clear that *some* coverage follows routinized dismissive patterns, this might not be that damaging to protests as long as substantive information about causes and grievances are still available elsewhere.
Studies following this logic suggest that the delegitimising aspects of the *protest paradigm* might occur in reporting more conditionally [@kilgoHierarchy2019; @woutersReportingDemonstrationsEpisodic2015], and less frequently over time [@gruberTrouble2023; @araizaHandsDontShoot2016; @kyriakidouIndignadosProtestsSpanish2017].

@@

It contributes to the literature in two ways 
Firstly, while this is not the first study to test factors which condition protest coverage[e.g., @boyleAdherenceProtestParadigm2012; @kilgoHierarchy2019; @leeTriggeringProtestParadigm2014], the set of variables presented here is novel and, to my knowledge, the most comprehensive so far.
The first section of the article discusses which mechanisms and protest features scholars in the *protest paradigm* tradition have hypothesised to drive coverage and distills them into seven event, outlet and time based factors.

@@

Results indicate that...

@@

This section interrogates the mechanisms driving protest coverage that have been brought up by scholars who studied the *protest paradigm*.
To structure the discussion, I divide the factors into event-level, outlet-level and time-related factors.^[Similar to how @shoemakerMediatingMessage21st2014 divide influences believed to shape the news into several levels ordered from micro to macro.]
It is important to recognize that the factors outlined below are rarely singular determinants of news coverage. Instead, a complex interplay of influences guides journalists in emphasizing specific aspects of a protest event over others. Some of these influences are consciously considered by journalists, while others remain implicit, as demonstrated by @tenenboimweinblattProducingProtestNews2014.

@@

Instead of ranking these types of protest according to the level of deviance, @kilgoHierarchy2019 suggest that "journalistic routines are subject to a *hierarchy of social struggle*, in which certain topics are given precedence and legitimacy, and others are delegitimised, trivialised, or ignored altogether" (p. 16).
They also add several new types to this list.
Specifically, they found that coverage of protest against anti-Black and anti-Indigenous racism, international protest, and anti-Trump protests use frames that delegitimise their issue far more often than other protest types.
For some issues, like health, the environment and immigration protests, the press used legitimising frames more often.
Similarly, @harlowWholeWorldWatching2020 coded 33 specific types of protests before summarising them into six categories.
They find that while not all protest types had a significant influence on the use of frames, some patterns emerged:
conservative protests, for example, were less likely to be covered in a riot frame but more likely to be covered in a spectacle frame, while socio-economic protests were more likely to be covered using the riot frame.
For a legitimising frame, dubbed debate frame, no significant relationship to a protest type was found.

Following @kilgoHierarchy2019, I expect that some goals are more likely to receive legitimising and some are more likely to receive delegitimising coverage, i.e. that a *hierarchy of social struggle* will emerge in UK reporting of protest.
I will use five broad categories for protest goals, ordered from most to least likely to receive critical coverage these are:
*political*, *anti-war*, *labour*, *police* and *social-issue* protests [@boyleAdherenceProtestParadigm2012; @boyleNewspapersProtestExamination2005; @boyleInfluenceLevelDeviance2004].
The hypotheses I will test are thus:

@@

Early studies often treated goals and tactics as two sides of the same coin.
However, @boyleMeasuringLevelDeviance2009 established that this is a shortcoming as both have distinct effects on the coverage of protest.
Specifically, they showed that when both are taken into account, extreme protest tactics facilitate critical coverage stronger than protest goals.
This insight had already been observed by one of the earliest studies of protest coverage:
@gitlinWholeWorldWatching1980 found that the amount of outright negative coverage, instead of trivialisation, increased as anti-war protests became more extreme.
@boyleAdherenceProtestParadigm2012, who additionally test the effects of protest location, protest type, and a group's goals in an international comparison of protest coverage, conclude that the tactics employed during a protest are the strongest determinant of delegitimising coverage.
Likewise, @WasowSeeding2020 concludes that nonviolent black-led protests during the civil rights movement led to more coverage of civil rights issues, while protester violence invoked language associated with disorder and calls for law and order.

@@

The relation between violent tactics and coverage has been known for a long time by both researchers and protesters [@gitlinWholeWorldWatching1980].
When tactics become more violent, coverage becomes less supportive of the cause and more supportive of repressive measures.
Nevertheless, protesters often use violent tactics because this will usually lead to (more) coverage [@boyleAdherenceProtestParadigm2012].
**H2a** and **H2b**, therefore, also speak to the question if this bargain makes sense for a movement.

@@

In the landmark study which defined the idea of a *protest paradigm*, @chanJournalisticParadigmsCivil1984 found that not all newspapers engage in delegitimising protests in the same way:
the left-leaning media were less hostile towards the protesting teachers and students in Hong Kong than centrist media, which were in turn less hostile than right-leaning journalists.
This factor has usually been overlooked in subsequent US-based studies, which might have assumed that ideological differences between major outlets are less important given the emphasis on objectivity in the country's journalistic culture.
More recent studies, however, have started again to employ newspaper ideology as an explaining factor for the level of delegitimising reporting different protests face [@leeTriggeringProtestParadigm2014; @shahinProtestingParadigmComparative2016; @ozMainstreamMediasCoverage2016].
UK media --- the focus of this study --- are known to be distinctly more partisan than outlets in the US and elsewhere [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004], which could increase the effect of newspaper ideology on protest reporting.
This leads to the expectation:

@@

Gavin [-@gavinPressTelevisionBritish2007, pp. 95-118] and @rosieAnarchistsWorldCup2009 add another factor related to newspaper outlets:
the difference between tabloid and broadsheet newspapers.
In general, it is expected that tabloid newspapers are more focused on human interest and entertainment [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004].
The sharp divide in reporting style between broadsheets newspapers (*The Daily Telegraph*, the *Financial Times*, *The Guardian*, *The Independent* and *The Times*), and tabloids (*Daily Express*, *Daily Mail*, the *Daily Mirror* and *The Sun*) is a well-known factor in the UK media system and should hence not be neglected.

@gavinPressTelevisionBritish2007, who assessed the May Day demonstration in London in 2000, concluded that tabloid outlets used the delegitimising elements from the *protest paradigm* more extensively and failed to mention any motivation or goals of the protest.
Broadsheet newspapers --- especially *The Guardian* --- apparently made an effort to explain why the protests were held.
In their qualitative study, @rosieAnarchistsWorldCup2009 find that tabloid newspapers were more extreme in their framing of peaceful G8 protesters as a fringe group of radical militants.
It is, therefore, only plausible to expect that the two types use different frames.
Following the work by @rosieAnarchistsWorldCup2009, the following expectations can be made:

@@

Two factors related to time are expected to have an influence on reporting about protests.
The first one is that protests that happened more recently receive more legitimising coverage.
In an earlier study, I showed that the level of legitimising framing in news about domestic protest has been steadily growing since the early 1990s [@gruberTrouble2023].
Several mechanisms have been suggested for this:
Protest has largely been normalised since the first seminal studies about movements in the 1960s have been conducted [@daltonDemocraticChallengesDemocratic2004; @sandersDynamicsProtestBritain2003; @norrisDemocraticPhoenixReinventing2002, pp. 188-212; @vanaelstWhoThatWo2001; @oliverPoliticalProcessesLocal2000].
Today, many different groups employ protest as a political communication tool, and many more people today have participated in a protest at least once in their lives [e.g., @borbathDifferentWorldsContention2020; @HutterChallenges2019].
Besides this, information about protesters' goals and grievances are more readily available on the internet [@humprechtMoreSameMarketplace2013; @posterWhatMatterInternet2001; @boydSocialNetworkSites2011], while journalists likely feel more pressure to paint a complete picture today as audiences have a sharply increased pool of alternative information sources [@vanaelstPoliticalCommunicationHighchoice2017].

@@

- `r hyp("H8a")`

@@

The *news framing cycle* intuitively makes sense, yet the proposed time frame might suffer from one problem:
Gottlieb’s [-@gottliebProtestNewsFraming2015] case, the Occupy Wall Street movement, was arguably special since protest stretched over a long period and media attention did not wane for months.
He, therefore, suggests that the second phase, when reporting shifts to more substantial coverage, started only in the third week after the protest started --- that is, at a time when the media will have already forgotten most protest events.

@@

I use media coverage of protests as the dependent variable.
Newspaper articles were employed instead of alternatives, such as online news, social media content or TV transcripts for two reasons.
Firstly, despite declining audiences, mainstream news media such as newspapers are still deemed the most important platforms for public debate and the creation of the public agenda [@rogstadTwitterJustRehashing2016].
Newspapers might have lost some of their intermedia agenda-setting power due to being slower to release breaking news, but the more in-depth writing produced by newspapers nevertheless forms a crucial share of what is redistributed via online news and social media and is key in initiating, amplifying and sustaining attention to an issue [@Langerwindrush2020; @harderIntermediaAgendaSetting2017; @rogstadTwitterJustRehashing2016].
Secondly, suitable data from UK newspapers were already archived and digitally available for the whole period of interest.

@tbl-outlets presents the selection of newspaper outlets, strategically chosen to contain variance on the two important factors:
ideology and type of newspaper in the UK
Additionally the selection was made to include the outlets with the highest circulation figures.
I retrieved articles from LexisNexis, utilizing the keywords "protest" and "demonstration" and several misspellings.
A pilot ensured the comprehensiveness of these keywords by testing a range of synonyms like "march", "rally", or "direct action", which, however, did not yield any additional relevant articles.
The articles underwent a processing phase using the R-package LexisNexisTools [@gruberlnt2021] which consisted of removing duplicate entries, items discussing non-domestic protests, and false positives.
To maintain theoretical integrity, I opted to limit our data to domestic protests, as studies have indicated notable differences in the coverage of non-domestic protests [@boyleAdherenceProtestParadigm2012].
This led to the final dataset, comprising 27,496 newspaper articles.

@@

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from @gruberTrouble2023.
In short, frames were identified by coding the frame elements found in Entman's [-@entman1993] definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected articles and combining the coded elements to frames using factor analysis.
This results in one binary variable per frame, which takes the value 1 if a frame is present and 0 otherwise.
Each article can contain several frames, or none.
After training supervised machine learning models on the manually coded random sample, all remaining articles were coded using these models.

@fig-frames show the seven identified frames and how often they were coded in news articles.
The majority of them were determined to delegitimise protest using many of the same patterns identified in the *protest paradigm* literature.
They portray protesters as a fringe group of freaks who disturb the social norms(Decay of Morals), highlight the violence caused by protesters (Troublemakers), suggest that protest is a bothersome interruption of everyday activities (Nuisance), or focus on police actions to reinstate public order (Law & Order).
Two legitimising frames, on the other hand, describe the main topic and cause of a protest event (Cause & Grievances) or highlight the accomplishments of the protesters for a just cause (Righteous Struggle).
The remaining frame, Police Violence can neither be said to be legitimising nor delegitimising, highlight the shortcomings of this dichotomy [@harlowNewParadigm2023].
The frame focuses on violence, which is usually assumed to distract from protest demands and hence delegitimising.
Contrary to most theoretical expectations, however, the frame portrays police as responsible for violence.
This could lead to more interest in and sympathy towards a protest goal.

@@

To merge the two datasets, I employed a two-staged strategy:
first, I selected all articles in the newspaper dataset that were published no earlier than seven days before an event started and no later than 30 days after it ended;
then I used the variables "Protest Locations", "Protest Group Identity" and "Protester Demands" to extract suitable keywords [see Appendix].
I then used keywords to filter the articles selected in the first stage:
articles were discarded if they did not feature at least one of the keywords.
This selection was manually validated and overlaps between events were separated by hand.
Overall, 5,639 articles could be matched with the 471 events from the database.
Most independent variables in the raw data were either enriched or recoded before the main analysis [see Appendix].

Using the variables listed in @tbl-variables, I fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectively constitute a panel dataset since observations must be seen as nested and not independent.
It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets.
However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016].
Therefore, I used newspaper-year as the Level 2 variable ($n = 180$).
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->
I assessed whether adding the newspaper-year level increases or decreases model fit.
To compare models, the Akaike information criterion (AIC) and the Bayesian information criterion (BIC) are calculated and compared using Analysis Of Variance (ANOVA) tests.
The lower the AIC and BIC, the better the model fits the data.
Both AIC and BIC are always considerably lower when including the second level into the modelling.
The p-values suggest that including the random effect improves all models significantly.
I examined if adding a newspaper-year level improves model fit.
The model comparisons were done using AIC, BIC, and ANOVA tests.
The inclusion of the second level in modelling significantly lowers both AIC and BIC, and the p-values indicate a significant improvement in all models.

@@

Discussion of results is divided into two steps.
I first refer to the frames aggregated into legitimising and delegitimising to test the hypotheses set up above.
This, however, neglects some of the details revealed through the individual seven frames.

@@

```{r}
#| results: asis
models <- readRDS("../misc/models.Rds")
model_logit_dele_mlm <- models$model_logit_dele_mlm
model_logit_le_mlm <- models$model_logit_le_mlm

@@

>>>>

---
title: "Whose voices are heard?"
subtitle: "Making sense of framing patterns in protest coverage"
abstract: |
  Existing research indicates that when mainstream news media report about demonstrations, protesters often face delegitimising coverage that focuses on, for example, clashes with the police or the appearance of protesters instead of their message.
  Recent studies have also found, however, that media treatment differs along several features of protest events.
  In this study, I examine the impact of these features through a large-scale automated framing analysis of UK protest events from 1992 to 2017, complemented by data from the Mass Mobilization Project (MMP).
  My findings highlight four main determinants influencing the use of legitimising or delegitimising framing in protest news coverage:
  (1.) Violence: Protests involving violence receive more delegitimising coverage and less legitimising coverage.
  (2.) Protest goals: The objectives of a protest significantly affect the nature of its coverage, although the relationships between frames and goals are complex. Notably, goals play a larger role in legitimising framing.
  (3.) Outlet characteristics: The type and ideology of a news outlet also impact coverage, though to a lesser extent than anticipated.
  (4.) Timing: More recent reports and those published a significant time after the event's onset are more likely to contain legitimising framing.

format:
  aft-pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
  aft-html: 
    citeproc: false
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
date: today
filters:
  - abstract-section
keywords:
    - "protest paradigm"
    - "media framing"
    - "social movements"
    - "news reporting"
    - "content analysis"
bibliography: references.bib
execute:
  echo: false
  eval: true
---

```{r setup}
#| eval: true
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  fig.align = "center",
  fig.path = "plots/"
)
source("../misc/functions.R")
library(tidyverse); theme_set(theme_minimal())
library(kableExtra)
library(texreg)
library(margins)
library(grid)
```

@@

Protest is a fundamental democratic tool for citizens to voice dissent about issues in society.
In modern democracies, however, citizens usually only hear about protest through mass media.
And this hearing consists not only of the audience becoming aware of a protest event, but ideally should also involve them receiving the information to *understand* the underlying issue and the causes for the public dissent.
However, most of the literature concerning protest coverage in news media points to a phenomenon referred to as *protest paradigm* to explain a recurring patterns in news about demonstrations:
the media focus on crime, clashes with police and the spectacle surrounding a protest or the protesters themselves and their looks, demographics or occupation, while neglecting to explain *why* protesters spent time and resources on organising and participating in a street demonstration in the first place.
The *protest paradigm* explains this by presuming a mental catalogue, supposedly common among journalists, which indicates to them "where to look (and where not to look), and informs them about what to discover" [@chanJournalisticParadigmsCivil1984, p. 187].
<!-- The *protest paradigm* is thus closely related to the literature on the sociology of newsmaking, which observed that news workers are "routinizing the unexpected" [@tuchmanMakingNewsDoing1973, p. 110] by creating practical and effective procedural rules, routines and mental catalogues of news story themes to cope with the onslaught of new events. -->

Most of the focus in previous studies has been spent on the dangers of this routinised reporting.
This is due to the normative core of the *protest paradigm*:
by regularly focusing on the action itself, rather than the sometimes more elusive and complicated message [@smithProtestAgendaBuilding2001] reporting shaped by the *protest paradigm* could lead to delegitimisation and marginalisation of protest events, protest movements, or protest as a political resource [@mcleodSocialControlSocial1999].
*Some* evidence for the *protest paradigm* has been found in most studies that analysed news coverage of protests over the years [for a systematic overview see @gruberTrouble2023].
However, more recent work has also asked *how common* delegitimising devices of the *protest paradigm* are, as nobody can deny that the press at least occasionally discusses the socio-political issues protesters try to rise awareness for --- hence *legitimising their actions*.
While it seems fairly clear that *some* coverage follows routinized dismissive patterns, this might not be that damaging to protests as long as substantive information about causes and grievances are still available elsewhere.
Studies following this logic suggest that the delegitimising aspects of the *protest paradigm* might occur in reporting more conditionally [@kilgoHierarchy2019; @woutersReportingDemonstrationsEpisodic2015], and less frequently over time [@gruberTrouble2023; @araizaHandsDontShoot2016; @kyriakidouIndignadosProtestsSpanish2017].

@@

It contributes to the literature in two ways.
Firstly, while this is not the first study to test factors which condition protest coverage [e.g., @boyleAdherenceProtestParadigm2012; @kilgoHierarchy2019; @leeTriggeringProtestParadigm2014], the set of variables presented here is novel and, to my knowledge, the most comprehensive so far.
The first section of the article discusses which mechanisms and protest features scholars in the *protest paradigm* tradition have hypothesised to drive coverage and distills them into seven event-, outlet- and time-based factors.

@@

Results indicate that the way protests are framed in news coverage is influenced by four primary factors:
**Violence**: incidences of violence in protests lead to more coverage that delegitimizes the protest and less that legitimizes it.
**Objectives of the protest**: the goals behind a protest play a role in shaping its media portrayal. The connection between these objectives and the framing is intricate, however, with protest goals having a more pronounced effect on legitimizing frames.
**Characteristics of the media outlet**: The nature and ideological leanings of a news organization also affect how protests are covered, although this influence is not as strong as initially thought.
**Timing of coverage**: reports that are more immediate after an event, and reports that are from more recent years tend to include more legitimizing frames in their coverage.

@@

To structure the review of previously identified factors that shape protest reporting, I interrogate factors on the event-level, outlet-level and time-related factors below.^[Similar to how @shoemakerMediatingMessage21st2014 divide influences believed to shape the news into several levels ordered from micro to macro.]
It is important to recognize, however, that the factors outlined below are rarely singular determinants of news coverage. Instead, a complex interplay of influences guides journalists in emphasizing specific aspects of a protest event over others. Some of these influences are consciously considered by journalists, while others remain implicit, as demonstrated by @tenenboimweinblattProducingProtestNews2014.

@@

Instead of ranking these types of protest according to the level of deviance, @kilgoHierarchy2019 suggest that "journalistic routines are subject to a *hierarchy of social struggle*, in which certain topics are given precedence and legitimacy, and others are delegitimised, trivialised, or ignored altogether" (p. 16).
They also add several new types to this list.
Specifically, they found that coverage of protest against anti-Black and anti-Indigenous racism, international protest, and anti-Trump protests use frames that delegitimise their issue far more often than other protest types.
For some issues, like health, the environment and immigration protests, the press used legitimising frames more often.
Similarly, @harlowWholeWorldWatching2020 tested the relation of different protest categories to the usage of several delegitimising and legitisimising frames.
They find that while not all protest types had a significant influence on the use of frames, some patterns emerged:
conservative protests, for example, were less likely to be covered in a riot frame but more likely to be covered in a spectacle frame, while socio-economic protests were more likely to be covered using the riot frame.
For a legitimising frame, dubbed debate frame, no significant relationship to a protest type was found.

Following the clarifications of the protest goal variable by @kilgoHierarchy2019 and @harlowWholeWorldWatching2020, I expect that some goals are more likely to receive legitimising and some are more likely to receive delegitimising coverage, i.e. that a *hierarchy of social struggle* will emerge in UK reporting of protest.
I will use five broad categories for protest goals, ordered from most to least likely to receive critical coverage these are:
*political*, *anti-war*, *labour*, *police* and *social-issue* protests [@boyleAdherenceProtestParadigm2012; @boyleNewspapersProtestExamination2005; @boyleInfluenceLevelDeviance2004].
The hypotheses I will test are thus:

@@

The relation between violent tactics and coverage has been known for a long time by both researchers and protesters [@gitlinWholeWorldWatching1980].
@boyleMeasuringLevelDeviance2009 showed that when both protest goals and tactics are taken into account, violent or disruptive tactics facilitate critical coverage stronger than protest goals.
<!-- This insight had already been observed by one of the earliest studies of protest coverage: -->
<!-- @gitlinWholeWorldWatching1980 found that the amount of outright negative coverage, instead of trivialisation, increased as anti-war protests became more extreme. -->
@boyleAdherenceProtestParadigm2012, who additionally test the effects of protest location, protest type, and a group's goals in an international comparison of protest coverage, conclude that the tactics employed during a protest are the strongest determinant of delegitimising coverage.
Likewise, @WasowSeeding2020 concludes that nonviolent black-led protests during the civil rights movement led to more coverage of civil rights issues, while protester violence invoked language associated with disorder and calls for law and order.

@@

<!-- The relation between violent tactics and coverage has been known for a long time by both researchers and protesters [@gitlinWholeWorldWatching1980]. -->
<!-- When tactics become more violent, coverage becomes less supportive of the cause and more supportive of repressive measures. -->
<!-- Nevertheless, protesters often use violent tactics because this will usually lead to (more) coverage [@boyleAdherenceProtestParadigm2012]. -->
<!-- **H2a** and **H2b**, therefore, also speak to the question if this bargain makes sense for a movement. -->

@@

In the landmark study which defined the idea of a *protest paradigm*, @chanJournalisticParadigmsCivil1984 found that not all newspapers engage in delegitimising protests in the same way:
the left-leaning media were less hostile towards the protesting teachers and students in Hong Kong than centrist media, which were in turn less hostile than right-leaning journalists.
<!-- This factor has usually been overlooked in subsequent US-based studies, which might have assumed that ideological differences between major outlets are less important given the emphasis on objectivity in the country's journalistic culture. -->
More recent studies, however, have started again to employ newspaper ideology as an explaining factor for the level of delegitimising reporting different protests face [@leeTriggeringProtestParadigm2014; @shahinProtestingParadigmComparative2016; @ozMainstreamMediasCoverage2016].
UK media --- the focus of this study --- are known to be distinctly more partisan than outlets in the US and elsewhere [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004], which could increase the effect of newspaper ideology on protest reporting.
This leads to the expectation:

@@

Gavin [-@gavinPressTelevisionBritish2007, pp. 95-118] and @rosieAnarchistsWorldCup2009 add another factor related to newspaper outlets:
the difference between tabloid and broadsheet newspapers.
In general, it is expected that tabloid newspapers are more focused on human interest and entertainment [@kuhnPoliticsMediaBritain2007; @hallinComparingMediaSystems2004].
And indeed, @gavinPressTelevisionBritish2007, who assessed the May Day demonstration in London in 2000, concluded that this focus translated to tabloid outlets using the delegitimising elements from the *protest paradigm* more extensively and failing to mention any motivation or goals of the protest.
He also found that broadsheet newspapers --- especially *The Guardian* --- made an effort to explain why the protests were held.
In their qualitative study, @rosieAnarchistsWorldCup2009 find that tabloid newspapers were more extreme in their framing of peaceful G8 protesters as a fringe group of radical militants.
<!-- The sharp divide in reporting style between broadsheets newspapers (*The Daily Telegraph*, the *Financial Times*, *The Guardian*, *The Independent* and *The Times*), and tabloids (*Daily Express*, *Daily Mail*, the *Daily Mirror* and *The Sun*) is a well-known factor in the UK media system and should hence not be neglected. -->
The sharp divide in reporting style between broadsheets newspapers and tabloids is a well-known factor in the UK media system and should hence not be neglected.

Following the work by @gavinPressTelevisionBritish2007 and @rosieAnarchistsWorldCup2009, the following expectations can be made:

@@

Two factors related to time are expected to have an influence on reporting about protests.
The first one is that protests that happened more recently receive more legitimising coverage.
In an earlier study, I showed that the level of legitimising framing in news about domestic protest has been steadily growing since the early 1990s [@gruberTrouble2023].
Several mechanisms have been suggested for this:
protest has largely been normalised since the first seminal studies about movements in the 1960s have been conducted [@daltonDemocraticChallengesDemocratic2004; @sandersDynamicsProtestBritain2003; @norrisDemocraticPhoenixReinventing2002, pp. 188-212; @vanaelstWhoThatWo2001; @oliverPoliticalProcessesLocal2000].
Today, many different groups employ street protest as a political communication tool, and many more people today have participated in a protest at least once in their lives [e.g., @borbathDifferentWorldsContention2020; @HutterChallenges2019].
Besides this, information about protesters' goals and grievances are more readily available on the internet [@humprechtMoreSameMarketplace2013; @posterWhatMatterInternet2001; @boydSocialNetworkSites2011], while journalists likely feel more pressure to paint a complete picture today as audiences have a sharply increased pool of alternative information sources [@vanaelstPoliticalCommunicationHighchoice2017].

@@

- `r hyp("H7")`

@@

The *news framing cycle* intuitively makes sense, yet the proposed time frame might suffer from one problem:
@gottliebProtestNewsFraming2015 suggests that the second phase, when reporting shifts to more substantial coverage, begins in the third week after the start --- that is, at a time when the media will have already forgotten most protest events.

@@

I use media coverage of protests as the dependent variable.
Newspaper articles were employed instead of alternatives, such as online news, social media content or TV transcripts for two reasons.
Firstly, despite declining audiences, mainstream news media such as newspapers are still deemed the most important platforms for public debate and the creation of the public agenda [@rogstadTwitterJustRehashing2016].
Newspapers might have lost some of their intermedia agenda-setting power, but the more in-depth writing produced by newspapers nevertheless forms a crucial share of what is redistributed via online news and social media and is key in initiating, amplifying and sustaining attention to an issue [@Langerwindrush2020; @harderIntermediaAgendaSetting2017; @rogstadTwitterJustRehashing2016].
Secondly, suitable data from UK newspapers were already archived and digitally available for the whole period of interest.

@tbl-outlets presents the selection of newspaper outlets, strategically chosen to contain variance on the two important factors:
ideology and type of newspaper in the UK
Additionally the selection was made to include the outlets with the highest circulation figures.
I retrieved articles from LexisNexis, utilizing the keywords "protest" and "demonstration" and several misspellings.
<!-- A pilot ensured the comprehensiveness of these keywords by testing a range of synonyms like "march", "rally", or "direct action", which, however, did not yield any additional relevant articles. -->
The articles underwent a processing phase using the R-package LexisNexisTools [@gruberlnt2021] which consisted of removing duplicate entries, items discussing non-domestic protests, and false positives.
To maintain theoretical integrity, I opted to limit the data to domestic protests, as studies have indicated notable differences in the coverage of non-domestic protests [@boyleAdherenceProtestParadigm2012].
This led to the final dataset, comprising 27,496 newspaper articles.

@@

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from @gruberTrouble2023.
In short, frames were identified by coding the frame elements found in Entman's [-@entman1993] definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected articles.
These elements were then combined to frames using factor analysis.
The result is one binary variable per frame, which takes the value 1 if a frame is present and 0 otherwise.
Each article can contain several frames, or none.
After training supervised machine learning models on the manually coded random sample, all remaining articles were coded using these models.

@fig-frames show the seven identified frames and how often they were coded in news articles.
The majority of them were determined to delegitimise protest using many of the same patterns identified in the *protest paradigm* literature.
They portray protesters as a fringe group of freaks who disturb the social norms (dubbed "decay of morals"), highlight the violence caused by protesters ("troublemakers"), suggest that protest is a bothersome interruption of everyday activities ("nuisance"), or focus on police actions to reinstate public order ("law & order").
Two legitimising frames, on the other hand, describe the main topic and cause of a protest event ("cause & grievances") or highlight the accomplishments of the protesters for a just cause ("righteous struggle").
The remaining frame, "Police Violence", can neither be said to be legitimising nor delegitimising, highlighting the shortcomings of this dichotomy [@harlowNewParadigm2023].
The frame focuses on violence, which is usually assumed to distract from protest demands and hence delegitimising.
Contrary to most theoretical expectations, however, the frame portrays police as responsible for violence.
This could lead to more interest in and sympathy towards a protest goal.

@@

I merged the two datasets by using the time frame of each event and some keywords (see Appendix).
Overall, 5,639 articles could be matched with the 471 events from the database.
Using the variables listed in @tbl-variables, I then fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectively constitute a panel dataset, since observations must be seen as nested and not independent.
<!-- I used newspaper-year as the Level 2 variable ($n = 180$), since the number of newspapers alone was considered too low for this design. -->
<!-- It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets. -->
<!-- However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016]. -->
<!-- Therefore, I used newspaper-year as the Level 2 variable ($n = 180$). -->
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->
<!-- I assessed whether adding the newspaper-year level increases or decreases model fit. -->
<!-- To compare models, the Akaike information criterion (AIC) and the Bayesian information criterion (BIC) are calculated and compared using Analysis Of Variance (ANOVA) tests. -->
<!-- The lower the AIC and BIC, the better the model fits the data. -->
<!-- Both AIC and BIC are always considerably lower when including the second level into the modelling. -->
<!-- The p-values suggest that including the random effect improves all models significantly. -->
<!-- I examined if adding a newspaper-year level improves model fit. -->
<!-- The model comparisons were done using AIC, BIC, and ANOVA tests. -->
The inclusion of the second level significantly lowers both AIC and BIC for all individual and grouped frame models.
<!-- , and the p-values indicate a significant improvement in all models. -->

@@

Discussion of results is divided into two steps.
I first refer to the frames aggregated into legitimising and delegitimising to test the hypotheses set up above.
This, however, neglects some of the details revealed through testing the relations to the individual seven frames, which is hence done in the following section.

@@

```{r}
#| results: asis
#| label: tab-regtwoframes
models <- readRDS("../misc/models.Rds")
model_logit_dele_mlm <- models$model_logit_dele_mlm
model_logit_le_mlm <- models$model_logit_le_mlm

@@

:::
